Reproducible Research: Peer Assessment 1
===========
```{r,echo=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE,cache=TRUE)
```

## Reading and processing data
downloading the zip file , unzipping it ,reading it using read.csv.The date column is in charcater data type casting it to date data type
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile='activity.zip')
unzip("activity.zip",exdir = "activity")
act = read.csv("activity/activity.csv")
act$date = as.Date(act$date,"%Y-%m-%d")
head(act)
summary(act)
pairs(act)
```

## Histogram of total number of steps taken each day
using dplyr package storing the number of total steps per day discarding the na values, using ggplot to plot the histogram of total number of steps taken each day
```{r}
library("ggplot2")
library(dplyr)
tem = act%>%group_by(date)%>%summarise(total_steps=sum(steps,na.rm = TRUE))
ggplot(tem,aes(x=total_steps))+geom_histogram(bins=5)
```

## Mean and Median number of steps taken each day
using summarise function in dplyr  package and finding the mean , median steps per day .
```{r}
tem%>%summarise(mean=mean(total_steps,na.rm=TRUE),median=median(total_steps,na.rm = TRUE))
```

Mean total steps per day is 9354.23 and the median total steps per day is 10395

## Time series plot of average number of steps taken
Using dplyr package to group the data by interval and finding the mean total steps per interval and plotting the time series of average steps.

```{r}
temp = act%>%group_by(interval)%>%summarise(avg_steps=mean(steps,na.rm = TRUE))
ggplot(temp,aes(x=interval,y=avg_steps))+geom_line()
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
interval_w_max_steps = arrange(temp,desc(avg_steps))$interval[1]
interval_w_max_steps
```
The 5-minute interval that, on average , contains the maximum number of steps is 835

## Strategy for dealing with missing values
The strategy I am going to use to deal with missing values is fill the missing values with mean values of steps per interval .
Create a new variable to store the new dataset with no NA values.
```{r}
totalna = sum(is.na(act))
totalna
```
There are total 2304 rows with missing values
```{r}
actnoNA = data.frame(act)
for(i in temp$interval){
    actnoNA[is.na(act$steps)&act$interval==i,"steps"]= temp[temp$interval==i,"avg_steps"]
}
head(actnoNA)
```

## Histogram of the total number of steps taken each day after missing values are imputed
Creating a histogram for total no of steps taken each day using the dataframe in which missing values are filled
```{r}
tem = actnoNA%>%group_by(date)%>%summarise(total_steps=sum(steps,na.rm = TRUE))
ggplot(tem,aes(x=total_steps))+geom_histogram(bins=5)
```
calculating the mean and median of steps per day 
```{r}
tem%>%summarise(mean=mean(total_steps,na.rm=TRUE),median=median(total_steps,na.rm = TRUE))
```

The mean and median vary of total steps per day after dealing with mssing values,the mean has increased and mean has become equal to median.
## Checking differences in activity patterns between weekdays and weekends
```{r}
wkend = c("Saturday","Sunday")
act = act%>%
    mutate(wktype=ifelse(weekdays(act$date) %in% wkend,"weekend","weekday"))
pl = act%>%group_by(wktype,interval)%>%summarise(avg_steps=mean(steps,na.rm = TRUE))
library(lattice)
xyplot(avg_steps ~ interval | wktype,pl, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```

From the plot we can see that there is a difference in average steps taken per day on weekdays and weekends.

